---
title: "Lab 05 - La Quinta is Spanish for next to Denny's, Pt. 2"
author: "John Bennett"
date: "Feb 22, 2022"
output: github_document
---

### Load packages and data

```{r load-packages, message = FALSE}
library(tidyverse) 
library(dsbox) 
```

```{r load-data, message = FALSE}
states <- read_csv("data/states.csv")
```

### Exercise 1
Q: How many Denny’s locations are there in Alaska?
A: There are 3 locations in AK.

```{r Ex1}
dennys_ak <- dennys %>%
  filter(state == "AK")
nrow(dennys_ak)
```

### Exercise 2
Q: How many La Quinta locations are there in Alaska?
A: There are 2 locations in AK.

```{r Ex2}
laquinta_ak <- laquinta %>%
  filter(state == "AK")
nrow(laquinta_ak)
```


### Exercise 3
Q: How many pairings are there between all Denny’s and all La Quinta locations in Alaska, i.e., how many distances do we need to calculate between the locations of these establishments in Alaska?
A: 11


```{r Ex3}
dennys_laquinta_ak <- full_join(dennys_ak, laquinta_ak, by = "state")
dennys_laquinta_ak
```

### Exercise 4.1
Q: How many observations are in the joined dennys_laquinta_ak data frame? 
A: There are 11 observations.

```{r Ex4}
ncol(dennys_laquinta_ak)
```

### Exercise 4.2
Q: What are the names of the variables in this data frame? 
A: Address.x, City.x, State.x, zip.x, longitude.x, latitude.x 


### Exercise 5 
Q: What function from the tidyverse do we use the add a new variable to a data frame while keeping the existing variables?
A: The mutate function will help us to add a new variable while keeping the existing variable. 

### Exercise 6 Create "distance" variable
Q: Calculate the distances between all pairs of Denny’s and La Quinta locations and save this variable as "distance". Make sure to save this variable in THE "dn_lq_ak" data frame so that you can use it later....
A: [Note: This is the point where I tried some creative problem solving...]

```{r Ex6DistanceVar}
dennys_laquinta_ak <- dennys_laquinta_ak %>%
  mutate(distancevar = case_when(longitude.x == -149.8767 ~ long1,
                                latitude.x == 61.1953 ~ lat1,
                                longitude.x == -149.8090 ~ long2,
                                latitude.x == 61.2097 ~ lat2,
                                longitude.x == -147.7600 ~ long3,
                                latitude.x == 64.8366 ~ lat3))
```


dennys_laquinta_ak <- dennys_laquinta_ak %>%
summary(dennys) %>%
glimpse(dennys)
```{r Ex6QuestDennys}
?dennys
```


```{r Ex6}

haversine <- function(long1, lat1, long2, lat2, round = 3) {
  # convert to radians
  long1 = long1 * pi / 180
  lat1  = lat1  * pi / 180
  long2 = long2 * pi / 180
  lat2  = lat2  * pi / 180
  
  R = 6371 # Earth mean radius in km
  
  a = sin((lat2 - lat1)/2)^2 + cos(lat1) * cos(lat2) * sin((long2 - long1)/2)^2
  d = R * 2 * asin(sqrt(a))
  
  return( round(d,round) ) # distance in km
}
```


Note: After manually entering the latitude and longitude below, the function did not appear to produce an actual distance in km... 
```{r Ex6Manual}
haversine <- function(long1, lat1, long2, lat2, round = 3) {
  # convert to radians
  long1 = -149.8767 * pi / 180
  lat1  = 61.1953  * pi / 180
  long2 = -149.8090 * pi / 180
  lat2  = 61.2097  * pi / 180
  
  R = 6371 # Earth mean radius in km
  
  a = sin((61.2097 - 61.1953)/2)^2 + cos(61.1953) * cos(61.2097) * sin((-149.8090 - -149.8767)/2)^2
  d = R * 2 * asin(sqrt(a))
  
  return( round(d,round) ) # distance in km
}
```

These are my attempts to research online and derive a formula to help create the "distance" variable. 
```{r Ex6Testing}
my_points <- matrix(c(81.25798, 73.81277,    # Create longitude/latitude matrix
                      14.91254, 18.18145),
                    nrow = 2)
colnames(my_points) <- c("longitude", "latitude")
rownames(my_points) <- c("point_1", "point_2")
my_points                                    # Print longitude/latitude matrix
```

```{r Ex6GeoPack}
install.packages("geosphere")                # Install & load geosphere
library("geosphere")
```

```{r Ex6DistHav}
my_dist <- distHaversine(my_points)          # Calculate Haversine distance
my_dist                                      # Print Haversine distance
```

### Exercise 7 
    [Note: Have to have "distance" variable, created in Ex6 above] Calculate the minimum distance between a Denny’s and La Quinta for each Denny’s location. To do so we group by Denny’s locations and calculate a new variable that stores the information for the minimum distance.

```{r Ex7}
dennys_laquinta_ak_mindist <- dennys_laquinta_ak %>%
  group_by(address.x) %>%
  summarize(closest = min(distance))
```


### Exercise 8 
[Note: I ran into a major brick wall on Exercise #6, which forced me to take a pause at this point and submit what I have, in the interest of time]

Describe the distribution of the distances Denny’s and the nearest La Quinta locations in Alaska. Also include an appripriate visualization and relevant summary statistics.

```{r Ex8}

```

### Exercise 9
Repeat the same analysis for North Carolina: (i) filter Denny’s and La Quinta Data Frames for NC, (ii) join these data frames to get a complete list of all possible pairings, (iii) calculate the distances between all possible pairings of Denny’s and La Quinta in NC, (iv) find the minimum distance between each Denny’s and La Quinta location, (v) visualize and describe the distribution of these shortest distances using appropriate summary statistics.

```{r Ex9}

```

### Exercise 10
Repeat the same analysis for Texas.


### Exercise 11
Repeat the same analysis for a state of your choosing, different than the ones we covered so far.


### Exercise 12
Among the states you examined, where is Mitch Hedberg’s joke most likely to hold true? Explain your reasoning.
